---
title: "Hossain_HW2_eco480"
author: "Husyn"
date: "March 13, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##3.1


## Loading in Data

```{r loading in .dta data}
library(readstata13)
data1 <- read.dta13("./cps92_12.dta")
```


a.

i.
```{r mean}
mean(data1$ahe[which(data1$year == 1992 | data1$year == 2012)])
```

ii.
```{r std dev}

sd(data1$ahe[which(data1$year == 1992 | data1$year == 2012)])

```

iii.
```{r 95CI}

library(gmodels)
ci(data1$ahe[which(data1$year == 1992 | data1$year == 2012)])

```
iv.
```{r dual}
a <- sqrt(((sd(data1$ahe[which(data1$year == 1992)])^2)/NROW(data1$ahe[which(data1$year == 1992)]))+((sd(data1$ahe[which(data1$year == 2012)])^2)/NROW(data1$ahe[which(data1$year == 2012)])))

abs(mean(data1$ahe[which(data1$year == 1992)]) - mean(data1$ahe[which(data1$year == 2012)])) + 1.96*a

abs(mean(data1$ahe[which(data1$year == 1992)]) - mean(data1$ahe[which(data1$year == 2012)])) - 1.96*a
```

b.
i.
```{r }
b <- (data1$ahe[which(data1$year == 1992)]/140.3)*229.6

mean(c(data1$ahe[which(data1$year == 2012)],b))

```

ii.
```{r}
sd(c((((data1$ahe[which(data1$year == 1992)])/140.3)*229.6), data1$ahe[which(data1$year == 1992)]))

```
iii.
```{r }
library(gmodels)
ci(c((((data1$ahe[which(data1$year == 1992)])/140.3)*229.6), data1$ahe[which(data1$year == 1992)]))


```
iv.
```{r }
c <- data1$ahe[which(data1$year == 1992)]
d <- (c/140.3) * 229.6
a <- sqrt(((sd(d)^2)/NROW(d)) + ((sd(data1$ahe[which(data1$year == 2012)])^2)/NROW(data1$ahe[which(data1$year == 2012)])) )

abs((mean(d) - mean(data1$ahe[which(data1$year == 2012)]))) + 1.96*a

abs((mean(d) - mean(data1$ahe[which(data1$year == 2012)]))) - 1.96*a

```
c.
You would use the results in b since inflation would distort purchasing power

d.
i.
```{r }
e <- data1$ahe[which(data1$year == 2012)]
ci(e)

```
ii.
```{r }
f <- data1$ahe[which(data1$year == 2012 & data1$bachelor == 1)]
ci(f)

```
iii.
```{r }
g <- sqrt((sd(e)^2 / NROW(e)) + (sd(f)^2 / NROW(f)))
abs(mean(e) - mean(f)) + g
abs(mean(e) - mean(f)) - g

```

e.
i.
```{r }
h <- (data1$ahe[which(data1$year == 1992)]/140.3)*229.6
ci(h)

```
ii.
```{r }
i <- (data1$ahe[which(data1$year == 1992 & data1$bachelor == 1)]/140.3)*229.6
ci(i)

```
iii.
```{r }
j <- sqrt((sd(h)^2 / NROW(h)) + (sd(i)^2 / NROW(i)))
abs(mean(e) - mean(f)) + j
abs(mean(e) - mean(f)) - j

```
f.
i.
H0 : mean AHE 1992 - mean AHE 2012  = 0
H1 : mean AHE 2012 - mean AHE 1992  > 0
```{r }
t.test(e,h, "greater")

```
Very low p value so we reject null, 2012 wages are higher than 1992.
ii.
H0 : mean AHE 1992 - mean AHE 2012  = 0
H1 : mean AHE 2012 - mean AHE 1992  > 0
```{r }
coll2012wg <- data1$ahe[which(data1$year == 2012 & data1$bachelor == 1)]
coll1992wg <- data1$ahe[which(data1$year == 1992 & data1$bachelor == 1)]

t.test(coll2012wg, coll1992wg, "greater")

```
The p value is very low, so we must reject null hypothesis. Real wages of college graduates did increase.

g.
``` {r }

library(psych)
data1992 <- data1[data1$year == 1992,]
data2012 <- data1[data1$year == 2012,]

describeBy(data1992,data1992$bachelor)

describeBy(data2012,data2012$bachelor)

```
The wages of college graduates increases more between 1992 and 2012 than high school graduates.

##3.2
##Loading in Data

```{r }
library(readstata13)
Sports <- readstata13::read.dta13("./Sportscards.dta")
```

a.
i.
Since 50% have good B, and all have the option to switch goods, and all have preference for good A, 50% of people will trade.

ii.

At least 50% of people would trade for the other good.

iii.

X% * n + (100 - X)% * n = 50%

b.
``` {r }
prcntTrd <- sum(Sports$trade)/nrow(Sports)

```
H0: %traded = 50%
H1: %traded < 50%

```{r }
t.test(x = Sports$trade, alternative = "less", mu = .5)

```

The p-value in the t test is very small, so we must reject null and determine that the fraction of 33.78% is significantly different from 50%

c.
```{r }
t.test(x = Sports$trade[Sports$dealer == 1], alternative = "less", mu = 0.5 )



```
The p value in this case is large, so we fail to reject the null hypothesis. We can conclude with 95% confidence that the population mean of traders who are dealers is not less than 0.5.

```{r}
t.test(x = Sports$trade[Sports$dealer == 0], alternative = "less", mu = 0.5 )


```
Here, the p value is very low, so we reject the null and conclude that the population mean of trads made by non dealers is less than 0.5. A stark contrast to the earlier findings with the dealers. This evidence is consistent with the hypothesis that the endowment effect disappears as traders gain more experience.










